---
title: "Chp10"
author: "Dan Feeney"
date: "4/28/2020"
output: html_document
---

## Constructing a logistic regression for whether a chimp pulls left or right with predictors of food and pro social behavior.
```{r echo = TRUE}
rm(list=ls())
library(rethinking)
data(chimpanzees)
d <- chimpanzees
d$treatment <- 1 + d$prosoc_left + 2*d$condition
xtabs( ~ treatment + prosoc_left + condition , d )
```

## L ~ Binom(1,Pi)
## logit(Pi) + alpha(chimp)(i) + beta(treatment)(i), with priors for a and b to be determined.
## This is also known as Li ~ Bernoulli(Pi). This model has 7 parameters (1 for each chimp) and 4 treatment parameters (each combo of social and food)
### Prior predictive simulations
```{r}
m11.1 <- quap( 
    alist(
    pulled_left ~ dbinom( 1 , p ) ,
    logit(p) <- a ,
    a ~ dnorm( 0 , 10 ) #with an alpha only model mean 0, sd of 10
  ) , data=d )
set.seed(5280)
prior <- extract.prior(m11.1, n = 1e4)

## converting logit values to probability with the inverse link function
p <- inv_logit( prior$a )
dens( p , adj=0.1 ) #Super bad prior- only values are at the extrememes! always or never happens
# a flat prior in the logit space is not a flat prior in probability space!
```
## adjust prior for alpha to be sd of 1.5
```{r}
m11.1 <- quap( 
    alist(
    pulled_left ~ dbinom( 1 , p ) ,
    logit(p) <- a ,
    a ~ dnorm( 0 , 1.5) #with an alpha only model mean 0, sd of 10
  ) , data=d )
set.seed(5280)
prior <- extract.prior(m11.1, n = 1e4)

## converting logit values to probability with the inverse link function
p <- inv_logit( prior$a )
dens( p , adj=0.1 ) #Super bad prior- only values are at the extrememes! always or never happens
# a flat prior in the logit space is not a flat prior in probability space!
```


## set the prior for the beta (intercept) terms with the same idea
```{r}
m11.2 <- quap(
    alist(
    pulled_left ~ dbinom( 1 , p ) ,
    logit(p) <- a + b[treatment] ,
    a ~ dnorm( 0 , 1.5 ),
    b[treatment] ~ dnorm( 0 , 10 )
    ) , data=d )
  set.seed(1999)

prior <- extract.prior( m11.2 , n=1e4 )
p <- sapply( 1:4 , function(k) inv_logit( prior$a + prior$b[,k] ) ) #extract the probabiity values from inverse logit

## we are interested the in the difference btetween treatments, so calculate difference for density plot
dens( abs( p[,1] - p[,2] ) , adj=0.1 )
```

## Again, a ridiculous prior with an SD for intercept of 10, so let's regularize it as shown below
```{r}
m11.3 <- quap( 
    alist(
    pulled_left ~ dbinom( 1 , p ) ,
    logit(p) <- a + b[treatment] ,
    a ~ dnorm( 0 , 1.5 ),
    b[treatment] ~ dnorm( 0 , 0.5 )
  ) , data=d )
set.seed(1999)
prior <- extract.prior( m11.3 , n=1e4 )
p <- sapply( 1:4 , function(k) inv_logit( prior$a + prior$b[,k] ) )
mean( abs( p[,1] - p[,2] ) )

```
## Now set up MCMC to run with ulam rather than quadratic approximation
```{r}
# prior trimmed data list 11.10
dat_list <- list(
  pulled_left = d$pulled_left,
  actor = d$actor,
  treatment = as.integer(d$treatment) )

m11.4 <- ulam(
  alist(
    pulled_left ~ dbinom( 1 , p ) ,
    logit(p) <- a[actor] + b[treatment] ,
    a[actor] ~ dnorm( 0 , 1.5 ),
    b[treatment] ~ dnorm( 0 , 0.5 )
  ) , data=dat_list , chains=4 , log_lik=TRUE )
precis( m11.4 , depth=2 )
```

### first 7 alpha terms are unique to each chimp and is their avg tendency to pull the left lever
### 
```{r}
post <- extract.samples(m11.4)
p_left <- inv_logit( post$a )
plot( precis( as.data.frame(p_left) ) , xlim=c(0,1) )

# 7 chimps on Y axis and tendency to pull 

labs <- c("R/N","L/N","R/P","L/P") 
plot( precis( m11.4 , depth=2 , pars="b" ) , labels=labs )

diffs <- list(
  db13 = post$b[,1] - post$b[,3],
  db24 = post$b[,2] - post$b[,4] )
plot( precis(diffs) ) #log odds of pulling the lever
```

```{r}
dat <- list( actor=rep(1:7,each=4) , treatment=rep(1:4,times=7) )
p_post <- link( m11.4 , data=dat )
p_mu <- apply( p_post , 2 , mean )

post <- extract.samples(m11.4)
mean( exp(post$b[,4]-post$b[,2]) ) #on average, the switch from 2 to 4 (adding a partner) multiplies the odds of pulling the lever by 0.92, an 8% reduction. 
```




