---
title: "Chp3"
author: "Dan Feeney"
date: "4/15/2020"
output: html_document
---
## Chapter 3 notes

### Sampling from a grid-approximate posterior and sample 10000 values with replacement

```{r echo = TRUE}
library(rethinking)
p_grid <- seq( from=0 , to=1 , length.out=1000 )
prob_p <- rep( 1 , 1000 )
prob_data <- dbinom( 6 , size=9 , prob=p_grid )
posterior <- prob_data * prob_p
posterior <- posterior / sum(posterior)

samples <- sample( p_grid , prob=posterior , size=1e4 , replace=TRUE )
plot(samples)
dens(samples)
```

## the following answers questions based on the posterior generated above.
### based on 6 water landings in 9 tosses, this is the posterior distribution and the code below
### shows how you would calcualte the % of the posterior laying above or below some threshold, which
### is an estimate of the true proportion of the world covered in water. `
```{r echo = TRUE}
# add up posterior probability where p < 0.5 
sum( posterior[ p_grid < 0.5 ] )

# how many samples are between 0.5 and 0.75?
sum( samples > 0.5 & samples < 0.75 ) / 1e4

# what is the boundary for lower 80% of posterior?
quantile(samples, 0.8)
# quantiles for an interval
quantile(samples, c(0.1,0.9))
```

### if you had to give a point estimate of the posterior. The maximum a posteriori estiamte (MAP) is 0.66
## for the example where we observed 6/9 'water' landings
```{r echo = TRUE}
p_grid[which.max(posterior)]
chainmode( samples , adj=0.01 )


sum( posterior*abs( 0.5 - p_grid ) )
loss <- sapply( p_grid , function(d) sum( posterior*abs( d - p_grid ) ) )
# loss contains a value for each possible decision of point estiamte to use based on a loss function
p_grid[ which.min(loss) ]
```

## Given two tosses and a probability of 0.7, what are the relative probs of observing 0, 1, or 2 water
```{r echo = TRUE}
rbinom( 1 , size=2 , prob=0.7 )
plot(rbinom( 10 , size=2 , prob=0.7 ))
#Generate 10000 to verify how often each value of (0,1,2) appear in proportion to its liklihood
dummy_w <- rbinom( 1e5 , size=2 , prob=0.7 )
table(dummy_w)/1e5


#Reminder from previous chapters. Simulate probability of seeing 0, 1, or 2 'waters' in globe tossing with a prior p of 0.7
dbinom(0:2, size = 2, prob = 0.7) #says we have a 9, 42, and 49% of seeing 0, 1, or 2 W's comee up if it follows a binomial dist
#we can also sample from the binary distribution with: sampling 10 times of the 2 tosses of the globe and getting 0,1,or 2 W's
rbinom(10, size = 2, prob = 0.7)
dummyW <- rbinom(1e4, size = 2, prob = 0.7)
table(dummyW)/1e4 #divide number of observations for 0, 1 and 2 by total number of trials

```

```{r echo = true}
#HW practice
library(rethinking)
rm(list=ls())
birth1 <- c(1,0,0,0,1,1,0,1,0,1,0,0,1,1,0,1,1,0,0,0,1,0,0,0,1,0,
0,0,0,1,1,1,0,1,0,1,1,1,0,1,0,1,1,0,1,0,0,1,1,0,1,0,0,0,0,0,0,0,
1,1,0,1,0,0,1,0,0,0,1,0,0,1,1,1,1,0,1,0,1,1,1,1,1,0,0,1,0,1,1,0,
1,0,1,1,1,0,1,1,1,1)
birth2 <- c(0,1,0,1,0,1,1,1,0,0,1,1,1,1,1,0,0,1,1,1,0,0,1,1,1,0,
1,1,1,0,1,1,1,0,1,0,0,1,1,1,1,0,0,1,0,1,1,1,1,1,1,1,1,1,1,1,1,1,
1,1,1,0,1,1,0,1,1,0,1,1,1,0,0,0,0,0,0,1,0,0,0,1,1,0,0,1,0,0,1,1,
0,0,0,1,1,1,0,0,0,0)
noTrials <- length(birth1) + length(birth2)
noBoys <- sum(birth1)+sum(birth2)
#compute  posterior dis for being a boy (1 is boy)
p_grid <- seq( from=0 , to=1 , length.out=noTrials )
prob_p <- rep( 1 , noTrials )
prob_data <- dbinom( noBoys , size=noTrials , prob=p_grid )
posterior <- prob_data * prob_p
posterior <- posterior / sum(posterior)
#Sample from posterior over the grid of number 0,1
samples <- sample( p_grid , prob=posterior , size=1e4 , replace=TRUE )
plot(samples)
dens(samples)
PI(samples)
HPDI(samples)
p_grid[which.max(posterior)]
# Simulate 10,000 replicates of 200 births and compare amount of boy births to posterior. 
simBirths <- rbinom(1e4, size = 200, prob = 0.55) 
dens(simBirths) #X axis is number of births in the replicate of 200
```
