---
title: "Chp3"
author: "Dan Feeney"
date: "4/15/2020"
output: html_document
---
## Chapter 3 notes

### Sampling from a grid-approximate posterior and sample 10000 values with replacement

```{r echo = TRUE}
library(rethinking)
p_grid <- seq( from=0 , to=1 , length.out=1000 )
prob_p <- rep( 1 , 1000 )
prob_data <- dbinom( 6 , size=9 , prob=p_grid )
posterior <- prob_data * prob_p
posterior <- posterior / sum(posterior)

samples <- sample( p_grid , prob=posterior , size=1e4 , replace=TRUE )
plot(samples)
dens(samples)
```

## the following answers questions based on the posterior generated above.
### based on 6 water landings in 9 tosses, this is the posterior distribution and the code below
### shows how you would calcualte the % of the posterior laying above or below some threshold, which
### is an estimate of the true proportion of the world covered in water. `
```{r echo = TRUE}
# add up posterior probability where p < 0.5 
sum( posterior[ p_grid < 0.5 ] )

# how many samples are between 0.5 and 0.75?
sum( samples > 0.5 & samples < 0.75 ) / 1e4

# what is the boundary for lower 80% of posterior?
quantile(samples, 0.8)
# quantiles for an interval
quantile(samples, c(0.1,0.9))
```

### if you had to give a point estimate of the posterior. The maximum a posteriori estiamte (MAP) is 0.66
## for the example where we observed 6/9 'water' landings
```{r echo = TRUE}
p_grid[which.max(posterior)]
chainmode( samples , adj=0.01 )


sum( posterior*abs( 0.5 - p_grid ) )
loss <- sapply( p_grid , function(d) sum( posterior*abs( d - p_grid ) ) )
# loss contains a value for each possible decision of point estiamte to use based on a loss function
p_grid[ which.min(loss) ]
```

## Given two tosses and a probability of 0.7, what are the relative probs of observing 0, 1, or 2 water
```{r echo = TRUE}
rbinom( 1 , size=2 , prob=0.7 )
plot(rbinom( 10 , size=2 , prob=0.7 ))
#Generate 10000 to verify how often each value of (0,1,2) appear in proportion to its liklihood
dummy_w <- rbinom( 1e5 , size=2 , prob=0.7 )
table(dummy_w)/1e5
```