---
title: "Chp3"
author: "Dan Feeney"
date: "4/15/2020"
output: html_document
---
## Chapter 3 notes

### Sampling from a grid-approximate posterior and sample 10000 values with replacement

```{r echo = TRUE}
library(rethinking)
p_grid <- seq( from=0 , to=1 , length.out=1000 )
prob_p <- rep( 1 , 1000 )
prob_data <- dbinom( 6 , size=9 , prob=p_grid )
posterior <- prob_data * prob_p
posterior <- posterior / sum(posterior)

samples <- sample( p_grid , prob=posterior , size=1e4 , replace=TRUE )
plot(samples)
dens(samples)
```

## the following answers questions based on the posterior generated above.
### based on 6 water landings in 9 tosses, this is the posterior distribution and the code below
### shows how you would calcualte the % of the posterior laying above or below some threshold, which
### is an estimate of the true proportion of the world covered in water. 
```{r echo = TRUE}
# add up posterior probability where p < 0.5 
sum( posterior[ p_grid < 0.5 ] )

# how many samples are between 0.5 and 0.75?
sum( samples > 0.5 & samples < 0.75 ) / 1e4

# what is the boundary for lower 80% of posterior?
quantile(samples, 0.8)
# quantiles for an interval
quantile(samples, c(0.1,0.9))
```

