---
title: "Chp8"
author: "Dan Feeney"
date: "4/21/2020"
output: html_document
---

# Interaction effects
```{r echo = TRUE}
rm(list=ls())
library(rethinking)
data(rugged)

#fitting a linear model between geographic ruggedness and GDP
d <- rugged
# make log version of outcome
d$log_gdp <- log( d$rgdppc_2000 )
# extract countries with GDP data
dd <- d[ complete.cases(d$rgdppc_2000) , ]
# rescale variables
dd$log_gdp_std <- dd$log_gdp / mean(dd$log_gdp)
dd$rugged_std <- dd$rugged / max(dd$rugged)

#ruggedness is scaled to its max so 0 is not rugged at all and 1 is the most rugged. GDP is scaled log GDP/avg value, so 1 is average 0.8 is less than avg and 1.2 is more than avg
```

```{r echo = TRUE}
m8.1 <- quap( 
  alist(
    log_gdp_std ~ dnorm( mu , sigma ) ,
    mu <- a + b*( rugged_std - 0.215 ) , #0.215 is average ruggedness
    a ~ dnorm( 1 , 1 ) , 
    b ~ dnorm( 0 , 1 ) ,
    sigma ~ dexp( 1 )
  ) , data=dd )

# lets do some prior simulations
set.seed(7)
prior <- extract.prior( m8.1 )

# set up the plot dimensions
plot( NULL , xlim=c(0,1) , ylim=c(0.5,1.5) ,
xlab="ruggedness" , ylab="log GDP" )
abline( h=min(dd$log_gdp_std) , lty=2 )
abline( h=max(dd$log_gdp_std) , lty=2 )

# draw 50 lines from the prior
rugged_seq <- seq( from=-0.1 , to=1.1 , length.out=30 )
mu <- link( m8.1 , post=prior , data=data.frame(rugged_std=rugged_seq) )
for ( i in 1:50 ) lines( rugged_seq , mu[i,] , col=col.alpha("black",0.3) )

## this is an implausible prior because it includes highly unrealistic values

# narrow the SD for alpha and beta and you get a more reasonbale prior
m8.1 <- quap(
  alist(
    log_gdp_std ~ dnorm( mu , sigma ) ,
    mu <- a + b*( rugged_std - 0.215 ) ,
    a ~ dnorm( 1 , 0.1 ) ,
    b ~ dnorm( 0 , 0.3 ) ,
    sigma ~ dexp(1)
  ) , data=dd )
prior <- extract.prior(m8.1)
# set up the plot dimensions
plot( NULL , xlim=c(0,1) , ylim=c(0.5,1.5) ,
xlab="ruggedness" , ylab="log GDP" )
abline( h=min(dd$log_gdp_std) , lty=2 )
abline( h=max(dd$log_gdp_std) , lty=2 )

# draw 50 lines from the prior
rugged_seq <- seq( from=-0.1 , to=1.1 , length.out=30 )
mu <- link( m8.1 , post=prior , data=data.frame(rugged_std=rugged_seq) )
for ( i in 1:50 ) lines( rugged_seq , mu[i,] , col=col.alpha("black",0.3) )

precis(m8.1)
```

## Most people fit this model by adding another term to linear model mui = a + b(ri-rhat) + lambdaAi, where Ai is an indicator variable for AFrican continent or not, but then we need a prior for lambda, which tells the model AFrica is more uncertain.
## instead, we should fit an intercept for Africa and an intercept for not AFrica. Mui = alpha(cid)[i] + b(ri-rhat) where (cid) is a continent index variable of 1 for AFtica and 2 otherwise. There are two alpha (intercept) parameters, one for each unique index value.

```{r echo = TRUE}
# make variable to index Africa (1) or not (2)
dd$cid <- ifelse( dd$cont_africa==1 , 1 , 2 )

m8.2 <- quap(
    alist(
      log_gdp_std ~ dnorm( mu , sigma ) ,
      mu <- a[cid] + b*( rugged_std - 0.215 ) ,
      a[cid] ~ dnorm( 1 , 0.1 ) , #this is where we specify the [cid] variable
      b ~ dnorm( 0 , 0.3 ) ,
      sigma ~ dexp( 1 )
    ) , data=dd )
compare(m8.1, m8.2) #Differnce of 64 in WIAC, so the continent variable is picking up a good bit of variance
precis(m8.2, depth = 2) #The a[1] is the intercept for African nations and seems reliably lower than that of non African
#Posterior contrast between the two alphas is below 0
post <- extract.samples(m8.2) 
diff_a1_a2 <- post$a[,1] - post$a[,2]
PI( diff_a1_a2 )
```

## Sample from posterior predictions of both (using link). Despite the predictive superiority of adding in a second alpha, the different slopes are not accounted for
```{r echo = TRUE}
rugged.seq <- seq( from=-0.1 , to=1.1 , length.out=30 ) 
# compute mu over samples, fixing cid=2
mu.NotAfrica <- link( m8.2 ,
  data=data.frame( cid=2 , rugged_std=rugged.seq ) )
# compute mu over samples, fixing cid=1
mu.Africa <- link( m8.2 ,
  data=data.frame( cid=1 , rugged_std=rugged.seq ) )
# summarize to means and intervals
mu.NotAfrica_mu <- apply( mu.NotAfrica , 2 , mean )
mu.NotAfrica_ci <- apply( mu.NotAfrica , 2 , PI , prob=0.97 )
mu.Africa_mu <- apply( mu.Africa , 2 , mean )
mu.Africa_ci <- apply( mu.Africa , 2 , PI , prob=0.97 )

## This only shows the intercept is lower for African nations, but does not alter the slope
```

## we need to add an interaction effect! Now MUi = ALPHA(cid)[i] + BETA[cid][i](ri - rhat). Again, we will use 1,2 for AFrica/not instead of fitting an extra term in the 0/1 common example (which adds uncertainty to the variable associated with 1)

```{r echo  = TRUE}
m8.3 <- quap(
  alist(
    log_gdp_std ~ dnorm( mu , sigma ) ,
    mu <- a[cid] + b[cid]*( rugged_std - 0.215 ) ,
    a[cid] ~ dnorm( 1 , 0.1 ) ,
    b[cid] ~ dnorm( 0 , 0.3 ) ,
    sigma ~ dexp( 1 )
  ) , data=dd )
precis(m8.3, depth = 3) #shows there are not only different intercepts, but different slopes for AFrican nations
compare( m8.1 , m8.2 , m8.3 , func=PSIS ) #significant portion of weight given to model 8.3
plot( PSIS( m8.3 , pointwise=TRUE )$k ) #There are some high Pareto K values 
```

## plotting the interaction: 2 plots: first show AFrican nations,overlay posterior mean and regression line
```{r echo = TRUE}
# plot Africa - cid=1 8.17
d.A1 <- dd[ dd$cid==1 , ]
plot( d.A1$rugged_std , d.A1$log_gdp_std , pch=16 , col=rangi2 ,
  xlab="ruggedness (standardized)" , ylab="log GDP (as proportion of mean)" ,
  xlim=c(0,1) )
# extract estimaes from posterior for 97% CI and plot with lines and shading
mu <- link( m8.3 , data=data.frame( cid=1 , rugged_std=rugged_seq ) )
mu_mean <- apply( mu , 2 , mean )
mu_ci <- apply( mu , 2 , PI , prob=0.97 )
lines( rugged_seq , mu_mean , lwd=2 )
shade( mu_ci , rugged_seq , col=col.alpha(rangi2,0.3) )
mtext("African nations")

# plot non-Africa - cid=2
d.A0 <- dd[ dd$cid==2 , ]
plot( d.A0$rugged_std , d.A0$log_gdp_std , pch=1 , col="black" ,
  xlab="ruggedness (standardized)" , ylab="log GDP (as proportion of mean)" ,
  xlim=c(0,1) )
mu <- link( m8.3 , data=data.frame( cid=2 , rugged_std=rugged_seq ) )
mu_mean <- apply( mu , 2 , mean )
mu_ci <- apply( mu , 2 , PI , prob=0.97 )
lines( rugged_seq , mu_mean , lwd=2 )
shade( mu_ci , rugged_seq )
mtext("Non-African nations")
```

## Continuous interactions
```{r echo = TRUE}
library(rethinking)
data(tulips)
d <- tulips
str(d)
#blooms will be outcome and water and shade are predictors

# center variables
d$blooms_std <- d$blooms / max(d$blooms)
d$water_cent <- d$water - mean(d$water)
d$shade_cent <- d$shade - mean(d$shade)
# Three parameters: alpha ~ normal(0.5,1) means when water and shade are at mean values, the bloomds should be half way to max bloom
# The two slopes are centered on 0, implying no information about the directionality, but we know there should be some direction to these relations
a <- rnorm( 1e4 , 0.5 , 1 ) 
sum( a < 0 | a > 1 ) / length( a )
#The prior SDs were all set to 1. The intercept for a must be >0 and < 1, but this prior assigns 63% of mass outside that range
# Try a smaller SD
a <- rnorm(1e4, 0.5, 0.25)
sum(a < 0 | a > 1)/length(a) #now only 5% is outside the range.
# now for slopes:
#The range of water and shade is 2 (for both, from -1 to 1). To go from blooms of 0 to 1, we need a slope of 0.5 from either variable. So a SD of 0.25 to both, then 95% of prior between -0.5 and +0.5, so either variable could (but unlikely) account for the entire range

m8.4 <- quap( 
    alist(
      blooms_std ~ dnorm( mu , sigma ) ,
      mu <- a + bw*water_cent + bs*shade_cent ,
      a ~ dnorm( 0.5 , 0.25 ) ,
      bw ~ dnorm( 0 , 0.25 ) ,
      bs ~ dnorm( 0 , 0.25 ) ,
      sigma ~ dexp( 1 )
    ) , data=d )
```


## building an interactio model: MU must be impacted by shade & water AND if you change one variable, it impacts the other. For example, if water is low, decreasing shade (increased light) can't help as much as when water is high. We want Betas to depend on the value of the other variable. 
### the model is MUi = ALPHA + LAMBDAw,i*Wi + BETASi, where LAMBDAwi is BETAw + BETAws + Si. We need a prior for the new Bws interaction parameter. Imagine the strongest plasuible interaction where enough shade makes water 0. If we set Si = 1, then the interaction should be equal and opposite to the main effect

```{r echo = TRUE}
m8.5 <- quap( 
    alist(
      blooms_std ~ dnorm( mu , sigma ) ,
      mu <- a + bw*water_cent + bs*shade_cent + bws*water_cent*shade_cent , #betas for water and shade plus an interaction
      a ~ dnorm( 0.5 , 0.25 ) , #same alpha as before
      bw ~ dnorm( 0 , 0.25 ) , #same slopes as before
      bs ~ dnorm( 0 , 0.25 ) ,
      bws ~ dnorm( 0 , 0.25 ) , #interactoin is equal and opposite to the slopes
      sigma ~ dexp( 1 )
    ) , data=d )
precis(m8.5, depth = 2)
```


## plotting interactions
### Triptych plots: 3 plots showing relation between water and blooms for a different value of shade. We can use -1,0,1 for three values of shade, but you can use a low, medium, and high value

```{r echo = TRUE}
#using the non interaction model
par(mfrow=c(1,3)) # 3 plots in 1 row
for ( s in -1:1 ) {
idx <- which( d$shade_cent==s )
plot( d$water_cent[idx] , d$blooms_std[idx] , xlim=c(-1,1) , ylim=c(0,1) ,
xlab="water" , ylab="blooms" , pch=16 , col=rangi2 )
mu <- link( m8.4 , data=data.frame( shade_cent=s , water_cent=-1:1 ) )
for ( i in 1:20 ) lines( -1:1 , mu[i,] , col=col.alpha("black",0.3) )
}

## using model 8.5 with the interaction, the slopes (the helpfulness of water) changes with levels of shade. With max shade, water does not really do much for a plant blooming. 
par(mfrow=c(1,3)) # 3 plots in 1 row
for ( s in -1:1 ) {
idx <- which( d$shade_cent==s )
plot( d$water_cent[idx] , d$blooms_std[idx] , xlim=c(-1,1) , ylim=c(0,1) ,
xlab="water" , ylab="blooms" , pch=16 , col=rangi2 )
mu <- link( m8.5 , data=data.frame( shade_cent=s , water_cent=-1:1 ) )
for ( i in 1:20 ) lines( -1:1 , mu[i,] , col=col.alpha("black",0.3) )
}
```